Predicting children's and adults' preferences in physical interactions via physics simulation

### 2021-05-10 paper/paper2021.Rmd
cogsci revisions

author-information: > 
    \author{{\large \bf George Kachergis (kachergis@stanford.edu)} \and {\large \bf Sama Radwan} \and {\large \bf Bria Long} \\ Department of Psychology, Stanford University \\ Palo Alto, CA 94305 USA
    \AND {\large \bf Judith Fan (jefan@ucsd.edu)} \\ Department of Psychology, University of San Diego \\ Madison, WI 53706 USA
    \AND {\large \bf Daniel Bear} \and {\large \bf Daniel Yamins} \\ Departments of Psychology & Computer Science, Stanford University \\ Palo Alto, CA 94305 USA
    \AND {\large \bf Michael C.~Frank} \\ Department of Psychology, Stanford University \\ Palo Alto, CA 94305 USA}

@@

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(here)
require(tidyverse)
require(tidyboot)
library(apa) # for generating APA-style text to report statistical tests
library(ggpubr)
require(gridExtra)
require(kableExtra)
```

@@

```{r human-vs-model-table, results="asis"}
# now we just have a table of selected features that were moderately correlated with people's choices
load(here("models/model_preds_beta1.RData"))

@@

kable(mod_tab, "latex", booktabs = T, escape=F, digits = 2, caption="Correlations ($r$) of model and human responses.") %>%
  column_spec(column = 2:ncol(mod_tab), width=c("1cm","1cm","1.3cm","1.3cm","1.3cm","1.3cm")) %>%
  kable_styling(latex_options = "scale_down") 
```

@@

Given the high reliability of choices on this task, particularly for adults ($r = .95$, children $r = .61$), we tested how well heuristic features of the physical interactions could predict these choices.
We explored a wide range of features that we thought might be proportional to people's interest in particular object interactions (e.g., how far a dropped object might end up from the drop location), but not one of these features explained much of the trial-level variance in people's choices. 
<!-- Instead, we found a few features that were partially correlated with people's choices: for example, the inverse standard deviation of the target's final position was the best predictor of both children's and adults' choices on support trials. -->
<!-- For containment trials, the normed velocity of the drop object after first collision was most correlated with adults' choices and moderately with children's, but children's choices were also correlated with the maximum and mean distance traveled by the target. -->
On balance, although much systematic variation in this dataset is unaccounted for, it is somewhat encouraging that people's interest in physical interactions can be in part accounted for by simple heuristic features based on simulations of the same physical task we asked participants to merely imagine.

>>>>

author-information: > 
    \author{
    {\large \bf George Kachergis}$^1,\star$, {\large \bf Samaher Radwan}$^1$, {\large \bf Bria Long}$^1$,  
    {\large \bf Judith Fan}$^2$, \\
    $^{\star}$\texttt{kachergis@stanford.edu} \\
    {\large \bf Daniel Bear}$^{1,3}$, {\large \bf Daniel Yamins}$^{1,3}$, {\large \bf Michael C.~Frank}$^1$ \\ 
    $^1$Department of Psychology, Stanford University \\
    $^2$Department of Psychology, University of California at San Diego \\
    $^3$Department of Computer Science, Stanford University}

@@

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(here)
require(tidyverse)
require(tidyboot)
library(apa) # for generating APA-style text to report statistical tests
library(ggpubr)
require(gridExtra)
require(kableExtra)
require(papaja)
```

@@

```{r human-vs-model-table, results="asis"} 
# to make double-wide, need to modify tex file to \begin{table*} / \end{table*}
# now we just have a table of selected features that were moderately correlated with people's choices
load(here("models/model_preds_beta1.RData"))

@@

#kable(mod_tab, "latex", booktabs = T, escape=F, digits = 2, caption="Correlations ($r$) of model and human responses.") 
  #column_spec(column = 2:ncol(mod_tab), width=c("1cm","1cm","1.3cm","1.3cm","1.3cm","1.3cm")) %>%
  #kable_styling(latex_options = "scale_down") 
apa_table(mod_tab, format="latex", digits=2, escape=F, span_text_columns=T,
          align = c("l", rep("r", ncol(mod_tab)-1)), # font_size = "small", 
          caption="Correlations ($r$) of model and human responses.")
```

@@

Given the high reliability of choices on this task, particularly for adults ($r = .95$, children $r = .61$), we tested how well heuristic features of the physical interactions could predict these choices.
In light of findings that people often single out one of several possible causes as "the" cause in judgements of physical causation [e.g., @gerstenberg2020], we might expect to find single features that are predictive of people's choices.
We explored a wide range of features that we thought might be proportional to people's interest in particular object interactions (e.g., how far a dropped object might end up from the drop location), but not one of these features explained much of the trial-level variance in people's choices. 
Instead, we found a few features that were partially correlated with people's choices: for example, the best predictor of both children's and adult's choices on support trials was the inverse standard deviation of the target's final position, but the strength of this correlation was only moderate ($r \approx 0.4$).
For containment trials, the normed velocity of the drop object after first collision was most correlated with adults' choices and moderately with children's, but children's choices were also correlated with the maximum and mean distance traveled by the target.
On balance, although much systematic variation in this dataset is unaccounted for, it is somewhat encouraging that people's interest in physical interactions can be in part accounted for by simple heuristic features based on simulations of the same physical task we asked participants to merely imagine.
